---
title: "Help yourself: Assignment 7"
author:
  - name: "Michael McCarthy"
    url: https://github.com/mccarthy-m-g
    affiliation: "PSYC 615 Lab"
    affiliation_url: https://github.com/mccarthy-m-g/psyc-615-lab
repository_url: https://github.com/mccarthy-m-g/psyc-615-lab
output:
  distill::distill_article:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>", collapse = TRUE)
```

## Prerequisites

To access the datasets, help pages, and functions that we will use in this *help yourself*, load the following packages:

```{r prerequisites}
here::i_am("labs/09_trend-analysis/help-yourself_assignment-7.Rmd")

library(here)
library(tidyverse)
library(afex)
library(effectsize)
library(emmeans)
library(broom)
library(ggpubr)
```

As well as this Help Yourself's example data:

```{r example-data}
# Note this is the same data from last week
scores <- readr::read_csv(
  here::here("data", "help-yourself_a6.csv"),
  col_types = "fffd"
)

# However, we're going to filter it so it's just for a one-way within-subjects
# design.
scores <- dplyr::filter(scores, task == "math")
```

## Trend Analysis

There isn't much new here from what you've been doing already, so we're just going to show you the code beginning to end.

```{r trend-analysis}
# First fit your ANOVA model, as usual
scores_model <- afex::aov_car(
  score ~ level + Error(id/(level)),
  type = 3,
  data = scores
)

# Second construct your reference grid, as usual
scores_emm <- emmeans::emmeans(scores_model, specs = "level")

# Third, do your contrasts. Note that we set the method to "poly" here. This is
# how we test for different order trends using polynomial contrasts.
scores_emm_contrast <- emmeans::contrast(
  scores_emm,
  method = "poly",
  infer = TRUE,
  adjust = "none"
)

# Fourth, convert t statistics to F statistics, and get the effect size for
# each trend. Note that the numerator degrees of freedom for each contrast
# is 1 when you convert to an F statistic.
scores_emm_contrast <- broom::tidy(scores_emm_contrast)

dplyr::mutate(
  scores_emm_contrast,
  statistic = statistic^2,
  effectsize::F_to_omega2(statistic, df = 1, df_error = df)
)
```

Finally, for plotting you can use the same code from Help Yourself Assignment 6, making the relevant adjustments for a one-way within-subjects ANOVA (since there we did a plot for higher-order).
