---
title: "Help yourself: Assignment 5"
author:
  - name: "Michael McCarthy"
    url: https://github.com/mccarthy-m-g
    affiliation: "PSYC 615 Lab"
    affiliation_url: https://github.com/mccarthy-m-g/psyc-615-lab
repository_url: https://github.com/mccarthy-m-g/psyc-615-lab
output:
  distill::distill_article:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>", collapse = TRUE)
```

## Prerequisites

To access the datasets, help pages, and functions that we will use in this *help yourself*, load the following packages:

```{r prerequisites}
here::i_am("labs/07_ancova/help-yourself_assignment-5.Rmd")

library(here)
library(tidyverse)
library(broom)
library(car)
library(emmeans)
library(ggpubr)
```

As well as this Help Yourself's example data:

```{r example-data}
grades <- readr::read_csv(here::here("data", "help-yourself_a5.csv"))
```

## Assumptions

You will need a bit of new code for the one of the new assumptions this week, the other new assumptions just use ANOVAs so we won't cover those here.

### Covariate-dependent variable correlation

The `cor.test()` function can be used to test the correlation between the covariate and the dependent variable using the `~ covariate + dv` formula syntax.

```{r cor-test}
cor.test(
  ~ minutes_listening + final_exam_grade,
  method = "pearson",
  data = grades
)
```

### Other assumptions

Normality, homogeneity of variance, the one-way ANOVA between the independent variable and covariate, and homogeneity of regression slopes assumptions all use functions and syntax you have used previously for other assignments. So check back if you need a reminder!

## ANCOVA

ANCOVA uses the familiar `aov()` function you have used throughout this course, adding the covariate into the formula is all that's needed. We also specify the contrast coding for the group variable to use orthogonal contrasts, rather than the traditional dummy coding (treatment contrasts) that are used by default in the `aov()` function. This ensures we partition variance correctly.

```{r ancova}
grades_model <- aov(
  final_exam_grade ~ group + minutes_listening,
  contrasts = list(group = contr.sum),
  data = grades
)
```

However, the `aov()` function uses Type 1 sums of squares, which you do not want for ANCOVA. To calculate ANCOVA with Type 3 sums of squares you can use the `Anova()` function from the **car** package, which takes the model above as its input, then specify the type of sums of squares you want.

```{r car-anova}
grades_model_t3 <- car::Anova(grades_model, type = 3)

broom::tidy(grades_model_t3)
```

Alternatively, you can use the `aov_car()` function from the **afex** package. The only downside to this is that it doesn't support the `tidy()` function we've been using to tidy our model objects up with, unless we extract one of the tables stored in the model object. It's also a bit more verbose.

```{r aov-car}
# The ANOVA functions from the afex package require that your data has an ID
# column, which is used for the error term, so we're adding one here
grades <- dplyr::mutate(grades, student = paste0("p", 1:length(group)))

grades_model_car <- afex::aov_car(
  final_exam_grade ~ group + minutes_listening + Error(student),
  data = grades,
  type = 3,
  factorize = FALSE # This is needed when running an ANCOVA
)

broom::tidy(grades_model_car$Anova)
```

If you find the extraction annoying you could just write a custom function, then save it to a script in the R directory so you can source it as needed.

```{r tidy-afex}
# The tidy function is a generic function that allows you to supply your own
# tidy methods for objects of different classes. Here we're making a tidy
# method for the afex_aov class.
tidy.afex_aov <- function(x, ...) {
  broom::tidy(x[["Anova"]])
}

# Now this works
broom::tidy(grades_model_car)
```

Alternatively, you could just use the `summary()` function. This is the standard way to get summary output for a model object if you're working interactively. The reason we waited so long to show this is that up until now, the `broom::tidy()` function served the same purpose and also had the extra utility of putting our output in a data frame, allowing us to further use our output for other calculations.

```{r summary}
summary(grades_model_car)
```

### Effect size

When you calculate effect size you should do it on the object created by the `Anova()` function.

```{r omega-squared}
effectsize::omega_squared(grades_model_t3)
```

If you use the afex package you can just do it on the model object.

```{r omega-squared-car}
effectsize::omega_squared(grades_model_car)
```

## Follow-ups

Nothing new! The `emmeans()`, `contrasts()`, and `eff_size()` functions with their familiar syntax and arguments is all you need.

However, to extract the residual standard deviation from the model using the `sigma()` function and the residual degrees of freedom using the `df.residual()` function, you have to do something slightly different depending on the function you chose above for the ANCOVA model.

```{r sigma}
# aov
sigma(grades_model)
df.residual(grades_model)

# afex
sigma(grades_model_car$lm)
df.residual(grades_model_car$lm)
```

The `sigma()` and `df.residual()` functions are generics just like the `tidy()` function, so you could also write a custom function for these.

### Plotting emmeans

However, you will use the output of emmeans in a new way. In order to plot the estimated marginal means with the correct error bars you can tidy the emmeans object then use that data for plotting.

```{r tidy-emmeans}
# First get the emmeans since we didn't do any follow-up code above
grades_emm <- emmeans::emmeans(grades_model, specs = "group")

# Then tidy it so the results are in a data frame that can be sent to ggpubr
grades_emmeans <- tidy(grades_emm, conf.int = TRUE)
grades_emmeans
```

Once you have the data it can be plotted. Since the exact values that are needed for plotting are stored in the data frame rather than raw data, this will look a bit different than usual. There are two steps here: First is the `ggbarplot()` call. Notice that we aren't using the optional `add` argument to add error bars. We can't since there's only one value per group in our data frame, and the lower and upper for the error bars are actually stored in their own columns. To plot them we need to add a new geom to the plot using ggplot2 `+` syntax.

After doing this, the second step is to use the `geom_errorbar()` function (`?ggplot2::geom_errorbar`) from the ggplot2 package to add the confidence intervals to the plot. You can just copy the syntax below for now. If you want to learn more about how ggplot2 works there are some readings in Learning more.

```{r ggpubr}
ggpubr::ggbarplot(grades_emmeans, x = "group", y = "estimate") +
  ggplot2::geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.4)
```

## Learning more

### ggplot2

- Introduction to ggplot2 <https://r4ds.had.co.nz/data-visualisation.html>
