---
title: "Help yourself: Assignment 8"
author:
  - name: "Michael McCarthy"
    url: https://github.com/mccarthy-m-g
    affiliation: "PSYC 615 Lab"
    affiliation_url: https://github.com/mccarthy-m-g/psyc-615-lab
repository_url: https://github.com/mccarthy-m-g/psyc-615-lab
output:
  distill::distill_article:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>", collapse = TRUE)
```

## Prerequisites

To access the datasets, help pages, and functions that we will use in this *help yourself*, load the following packages:

```{r prerequisites}
here::i_am("labs/10_mixed-anova/help-yourself_assignment-8.Rmd")

library(here)
library(tidyverse)
library(ggpubr)

```

```{r}
drinks <- readr::read_csv(
  here::here("data", "help-yourself_a8.csv"),
  col_types = "fffd"
  )
```

## Mixed ANOVA

You can probably guess how to do a mixed ANOVA already given what you've learned so far. But here's the code anyways!

```{r mixed-anova}
drinks_model <- afex::aov_car(
  rating ~ drink * day  + Error(id/day),
  type = 3,
  data = drinks
)

summary(drinks_model)
```

That's it! We'll leave it up to you to figure out how to do assumptions, follow-ups, and visualizations. No new code is needed, so just look back to previous weeks and you'll be able to figure it out.

## Learning more R beyond this course

Thanks for sticking with R through the course! We hope you've enjoyed it and will continue to use it in the future. Given the time constraints of the course there's a lot we could not teach you about R programming, but hopefully you now feel confident to learn more on your own. There's a lot of high-quality free material you can use to learn more about R programming and using other statistical methods in R.

I (Michael) keep a list of useful books on the wiki for my blog under the Library tab, which you can check out if you want somewhere to start from: https://github.com/mccarthy-m-g/tidytales/wiki/. Most of the books are free, a few are not. From that list I would recommend starting with R for Data Science; you'll get a lot out of that book. I would also recommend reading Statistical Rethinking at some point. It's about Bayesian statistics, but even if you aren't interested in Bayes, the book still has a lot to offer. And if you are serious about using R in your work I would also recommend reading Happy Git and GitHub for the UseR. Git is a version control software for programming projects, GitHub is sort of like a cloud repository for any projects you do that use Git. I use Git and GitHub for everything I do, it's a bit obtuse at first, but once you get the hang of it it becomes a critical part of your workflow. It also makes collaboration with others a lot easier. You can even file issues or make contributions to other R packages through GitHub.

If you look at the Wrangle, Visualize, Model, and Communicate tabs on the wiki, you will also find lists of R packages that are useful for different things.

You can also check out the [Big Book of R](https://www.bigbookofr.com), which is basically just a giant glossary of R books that are out there, arranged by topic.
