---
title             : "Assignment 1---Assumptions and T-Tests."
shorttitle        : "Assignment 1"

params:
  firstname  : First
  lastname   : Last
  studentid  : 0123456789
  TAs        : "Kelsey Cnudde & Michael McCarthy"
  show_output: FALSE
  assignment : 1

output: papaja::apa6_docx
knit: ( function(input, ...) { source("../../R/knit.R"); knit_assignment(input, ...) })
---

```{r setup, include=FALSE}
# set up here
here::i_am("assignments/assignment-1/assignment-1.Rmd")

# load packages
library(here)
library(papaja)

# set default chunk options
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  echo = params$show_output,
  warning = params$show_output,
  message = params$show_output,
  fig.retina = 0.8, # figures are either vectors or 300 dpi diagrams
  dpi = 300,
  fig.width = 6,
  fig.asp = 0.618, # 1 / phi
  fig.show = "hold",
  cache.extra = knitr::rand_seed
)
```

# Question 1

*Because it’s fun (or rather, educational), you will be experimenting with making your own datasets. Please include your final data points under each question; no need for an APA formatted table, listing them is fine!*

## Question 1a

*Make a positively skewed distributed (i.e., NOT normal) set of 18 participants on a fictitious variable (think of a variable that makes sense with the numbers). Report your variable’s mean and SD. Report and interpret the skewness numerical measure and an appropriate statistical test of normality. (3 marks)*

Answers here.

\newpage

## Question 1b

*Now, create a normally distributed dataset (n = 18), with the same mean and standard deviation from 1a. (as close as possible, and cannot be more than 10% - 20% difference from either the mean or the SD). Report your mean and SD. Report and interpret the skewness numerical measure and an appropriate statistical test of normality. (3 marks)*

Answers here.

\newpage

# Question 2

*From D2L, please download and import `assignment-01_q2.csv` into R.*

*Background: Imagine you are a TA for an undergrad statistics course (or maybe you are now!) and your students just finished their first statistics exam of the year. You, a clever statistical student yourself, decided to record the following variables from the students: age, gender (1 = female, 2 = male), how many hours they study per week, their love of statistics (1 to 10 scale), and their heart rate both before the exam and after the exam (normal range is 60-100 for adults). You are now ready to investigate!*

```{r Q2-dataset}
# You can read the data in this chunk

```

## Question 2a

*Briefly describe why independence is important for analyses that assume it (1 mark). Give one potential limitation of why this sample (i.e., the students) might not be independent? (1 mark)*

Answers here.

\newpage

## Question 2b

*First things first: explore your data set as a whole. Do you see any “weirdness” in any of the continuous variables (i.e., age, study, love, heart rates)? Explain what the weirdness is (including participant #s) and justify why you think it is weird for each instance (i.e., with visuals, statistics, etc.). With this information, would you remove any participants/data points from the future analyses? Why or why not? (5 marks)*

*Hint: Read the variable descriptions above.* 

*Note 1: Please only include the relevant information here (we don’t need a novel). Just point out the glaring issues (you’ll see them we hope). Please don’t transform the data or anything drastic like that if there is skewness or something like that!*

*Note 2. If you say that certain participants/data should be removed from the analyses, please do that for the remainder of the questions!*

Answers here.

\newpage

## Question 2c

*You are interested in whether males are significantly lower in the number of hours they study each week compared to females (Hypothesis 1). Write an APA formatted results section exploring Hypothesis 1. (5 marks) This includes stating the appropriate test (no justification needed), quickly reporting on assumptions (we assume independence from here on so no need for that, and only report normality after removing outliers – no need to repeat normality tests from b.) and reporting the findings (provide and reference an APA formatted figure and reference whether the hypothesis was supported or not). Hint: When reporting findings, you need to include practical significance!*

Answers here.

\newpage

## Question 2d

*You are also interested in whether there is a statistically significant increase in heart rates from before to after the exam for all students (Hypothesis 2). Write an APA formatted results section for Hypothesis 2. (5 marks)*

Answers here.

\newpage

## Question 2e

*Another TA brags to you that they collected similar data when they were teaching the class and their students’ love of statistics had a mean of 5.17. They claim that your class’ love was significantly lower (i.e., you aren’t as awesome as them). Does this TA have the right to brag, or should they spend more time studying statistics themselves? Choose an appropriate test and report the findings that justify your response to this TA (no APA results section needed, including no figure, assumptions, or effect sizes needed). (2 marks)*

Answers here.

\newpage

# Bonus

*(.5 marks; not exceed 100% on assignment) When the two groups for an independent t-test do not have equal variance, what specific value is first adjusted to compensate for this in the analysis?*

Answers here.
