---
title: "Help yourself: Assignment 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prerequisites

To access the datasets, help pages, and functions that we will use in this *help yourself*, load the following packages:

```{r prerequisites}
here::i_am("labs/02_assumptions-ttests/help-yourself_assignment-1.Rmd")

library(here)
library(tidyverse)
library(ggpubr)
library(papaja)
```

If you run this code and get the error message "there is no package called 'tidyverse'" (or a different package name), you will need to first install it, then run `library()` once again.

```{r, eval=FALSE}
install.packages("tidyverse")
library(tidyverse)
```

You only need to install a package once, but you need to reload it every time you start a new session.

If you need to be explicit about where a function (or dataset) comes from, you can use the special form `package::function()`. For example, `ggplot2::ggplot()` tells R explicitly that you are using the `ggplot()` function from the **ggplot2** package.

Telling R explicitly where a function comes from is called *namespacing*. The code below will always be explicit about where functions from loaded packages come from so that it is easier for you to help yourself when you want to learn more about a given function. However, you do not need to namespace functions in your code for assignments so long as you have the prerequisite packages loaded in your session.

## Creating data sets

There are many ways to create data sets in R. If you want to create a data set where the code is laid out like a spreadsheet you can use the `tribble()` function from the **tibble** package. This is useful if you want to create a small data set with values you input manually. To get help using this function you can run `?tibble::tribble` in the R console.

```{r tribble}
simpsons <- tibble::tribble(
  ~simpson, ~age,
  "homer",  36,
  "marge",  34,
  "bart",   10,
  "lisa",    8,
  "maggie",  1
)

simpsons
```

Alternatively, if you want to create a data set where each column is a vector of values you can use the `tibble()` function from the tibble package. To get help using this function you can run `?tibble::tibble` in the R console.

```{r tibble}
# Here we recreate the data set from the code chunk above using the tibble
# function
simpson_name <- c("homer", "marge", "bart", "lisa", "maggie")
simpson_age  <- c(36, 34, 10, 8, 1)

simpsons <- tibble::tibble(simpson = simpson_name, age = simpson_age)

simpsons
```

### Simulating data

Creating data sets using the `tibble()` function is useful when you want to simulate data instead of entering it manually. R comes built in with a number of functions to simulate data with different distributions. You can see them all by running `?distributions` in the console. The most useful distribution to simulate in this course will be the normal distribution, which can be simulated with the `rnorm()` function. To get help using this function you can run `?rnorm` in the R console.

```{r simulate}
normal_distribution <- rnorm(n = 10, mean = 30, sd = 5)

normal_distribution
```

All the simulation functions in R output vectors of values, which means you can use them with the `tibble()` function to create simulated data sets.

```{r simulated-dataset}
# Here we construct a data set using the simulated normal distribution from the
# code chunk above, with an additional column containing the participant ID for
# each simulated value
simulated_dataset <- tibble::tibble(id = 1:10, value = normal_distribution)

simulated_dataset
```

## Wrangling data

It is often the case that the data you work with needs wrangling to make it fit for use. This could mean cleaning the data, but it could also mean adding new columns, removing rows, calculating new values, or transforming old values. The **dplyr** package has you covered for all of these actions and more. There are six main functions that will cover the majority of your data wrangling needs:

### Selecting

The `select()` function (`?dplyr::select`) allows you to select (and optionally rename) columns in a data frame.

```{r select}
dplyr::select(.data = simpsons, age)
```

### Filtering

The `filter()` function (`?dplyr::filter`) is used to subset a data frame, retaining all rows that satisfy your conditions.

```{r filter}
dplyr::filter(.data = simpsons, simpson == "bart")
```

### Mutating

The `mutate()` function (`?dplyr::mutate`) adds new columns and preserves existing ones

```{r mutate}
dplyr::mutate(.data = simpsons, age_in_2021 = age + 34)
```

### Grouping and summarizing

The `group_by()` function (`?dplyr::group_by`) allows you to organize observations into groups and do data operations by group. The `ungroup()` function (`?dplyr::ungroup`) removes grouping.

The `summarize()` function (`?dplyr::summarize`) allows you to compute summary statistics on a data set either as a whole or by group by pairing it with the `group_by()` function.

```{r groupby-summarize}
# First we need a variable to group by. We can create it using the mutate
# function.
simpsons <- dplyr::mutate(
  .data = simpsons,
  sex = c("male", "female", "male", "female", "female"),
  .before = age
)

# Then we group the data by sex
simpsons_grouped <- dplyr::group_by(.data = simpsons, sex)

# And finally we calculate some summary statistics using our grouping variable
simpsons_grouped <- dplyr::summarize(
  .data = simpsons_grouped,
  group_size = dplyr::n(),
  mean_age = mean(age),
  sd_age = sd(age)
)

simpsons_grouped
```

The `summarize()` function can also be used independently to get summary statistics for the whole sample.

```{r summarize}
dplyr::summarize(
  .data = simpsons,
  total_age = sum(age),
  median_age = median(age),
  mean_age = mean(age),
  sd_age = sd(age)
)
```

### Arranging

The `arrange()` function (`?dplyr::arrange`) orders the rows of a data frame by the values of selected columns.

```{r arrange-ascending}
dplyr::arrange(.data = simpsons, age)
```

You can use the the `desc()` function (`?dplyr::desc`) to order the rows in descending order.

```{r arrange-descending}
dplyr::arrange(.data = simpsons, dplyr::desc(age))
```

## File paths

Before you can read data you need to tell R where the data file is located. There are several ways to do this, some of them better than others.

The easiest---but also easiest to break---way is to write a character string with a path to the file, relative to your current working directory. You can check your current working directory by running `getwd()` in the console. If you do this, you should see that your current working directory is the `psyc-615-lab` directory

```{r string-path}
string_path <- "data/help-yourself_01.csv"
string_path
```

A better way is to use the `file.path()` function (`?file.path`), which constructs the path to a file from components in a platform-independent way. This avoids file paths breaking between different operating systems, since Windows uses backslashes and OSX and Linux use forward slashes. 

```{r function-path}
function_path <- file.path("data", "help-yourself_01.csv")
function_path
```

The best way is to use the `here()` function (`?here::here`) from the **here** package. The `here()` function will always locate the files relative to your project root, in this case the `psyc-615-lab` directory. This is more consistent than locating files relative to your current working directory, and encourages the use of [project-oriented workflows](https://r4ds.had.co.nz/workflow-projects.html). You can use it as a drop-in replacement for `file.path()`.

```{r here-path}
here_path <- here::here("data", "help-yourself_01.csv")
here_path
```

You should read the here package vignettes to learn more about using `here()` in general (`vignette("here", package = "here")`) and in R Markdown documents such as this one (`vignette("rmarkdown", package = "here")`). After reading, if you scroll to line 15 of this document you will note that the `here::i_am()` function has already been set up for you. But you may need to set it up on your own at other points in this course.

## Reading data

There are a number of functions you can use to read different data files in R. For this course there are only two you need to worry about. To read CSV data files you can use the `read_csv()` (`?readr::read_csv`) function from the **readr** package.

```{r read-csv}
csv_data <- readr::read_csv(here::here("data", "help-yourself_01.csv"))
csv_data
```

To read SPSS `.sav` files you can use the `read_sav()` function (`?haven::read_sav`) from the **haven** package.

```{r read-spss}
sav_data <- haven::read_sav(here::here("data", "help-yourself_01.sav"))
sav_data
```

## Visualizing data

R is a powerhouse for visualizing data, and makes it easy to create publication ready plots and tables. 

### Plots

The plots you will make in this course will be simple, but if you enjoy data visualization then diving deeper into R's plotting capabilities is a great investment (check out the *#tidytuesday* tag on Twitter if you want a taste). Data visualization in R is typically done with the **ggplot2** package, which is based on a grammar of graphics that allow you to assemble plots step by step to your exact specifications. A companion to ggplot2 is the **ggpubr** package, which simplifies the creation of plots common to scientific publications in APA style.

#### Histograms and box plots

Histograms are useful for visualizing the distribution of observations for a given variable. The `gghistogram()` function (`?ggpubr::gghistogram`) can be used to create a histogram. You can change the number and width of bins with the optional `bins` and `binwidth` arguments.

```{r gghistogram}
ggpubr::gghistogram(data = csv_data, x = "anxiety")
```

An alternative way to visualize the distribution of observations for a given variable is with boxplots. This can be useful if you want to emphasize different percentiles of the data. The `ggboxplot()` function (`?ggpubr::ggboxplot`) can be used to create a box plot.

```{r ggboxplot}
ggpubr::ggboxplot(data = csv_data, y = "anxiety")
```

#### QQ plots

```{r ggqqplot}
ggpubr::ggqqplot(data = csv_data, x = "anxiety")
```

#### Bar plots and line plots

Bar plots can be used to visualize group differences. The `ggbarplot()` function (`?ggpubr::ggbarplot`) can be used to create a bar plot. The optional `add` argument can be used to add descriptive statistics to the bars, such as means and confidence intervals.

```{r ggbarplot}
ggpubr::ggbarplot(data = csv_data, x = "group", y = "anxiety", add = "mean_ci")
```

An alternative to bar plots are line plots. These can be useful when visualizing changes in repeated measures data.

```{r ggline}
ggpubr::ggline(data = csv_data, x = "group", y = "anxiety", add = "mean_ci")
```

### Tables

R also has great packages for making tables. Unfortunately you will not get to play around with these in this course. Partly because you will only ever make very simple tables in this course, and partly because R's best table making packages do not support Microsoft Word documents at the moment.

The easiest way to output an APA style table in an R Markdown document is to use the `apa_table()` function (`?papaja::apa_table`) from the **papaja** package. This function takes a data frame and APA styles it when you knit your document, with an optional caption using the `caption` argument.

```{r papaja-table}
# If you run this code in the console the output will look like gibberish. You
# need to knit the document to see the proper result.
papaja::apa_table(simpsons, caption = "Table showing Simpsons ages")
```

However, you should note that the `apa_table()` function will not do all the APA styling work for you. Things such as capitalization, number of digits printed, and so forth must be either: (1) Handled beforehand in the data frame given to the `apa_table()` function; or (2) handled afterwords manually in Microsoft Word.

... maybe include how to wrangle for apa styling

<!-- Note to future TAs or students who are fine messing with development versions of R packages: The gt package will have word support in the future. As of Aug 31, 2021, there is a PR in development on Word support https://github.com/rstudio/gt/pull/811. Once this is complete gt will likely be a better option than papaja for APA tables. Here is a blogpost showing how to make APA-like tables with gt https://www.anthonyschmidt.co/post/2020-06-03-making-apa-tables-with-gt/. This could be used currently by creating a gt table then saving it as a png; it's not as good as a native table in Word but it's workable and would allow students to use gt's capabilities for formatting nice tables. -->

## Calculating descriptive statistics

mean, sd, 

## Determining normality

skewness, kurtosis, 

Shapiro-Wilk test
Kolmogorov-Smirnov test

Levene's test

## t-tests
