---
title: "Help yourself: Assignment 3"
author:
  - name: "Michael McCarthy"
    url: https://github.com/mccarthy-m-g
    affiliation: "PSYC 615 Lab"
    affiliation_url: https://github.com/mccarthy-m-g/psyc-615-lab
repository_url: https://github.com/mccarthy-m-g/psyc-615-lab
output:
  distill::distill_article:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>", collapse = TRUE)
```

## Prerequisites

To access the datasets, help pages, and functions that we will use in this *help yourself*, load the following packages:

```{r prerequisites}
here::i_am("labs/04_orthogonal-contrasts/help-yourself_assignment-3.Rmd")

library(here)
library(tidyverse)
library(ggpubr)
library(broom)
library(effectsize)
library(emmeans)
```

As well as this Help Yourself's example data:

```{r example-data}
birthdays <- readr::read_csv(here::here("data", "help-yourself_a2.csv"))
```

## Display math

You will need to show your calculations for demonstrating orthogonality in planned comparisons for this week's assignment. You can fill in the math equations below for that.

$$
\begin{aligned} \\
[...]\cdot[...] &= x_1*y_1 + x_2*y_2 + x_3*y_3 \\
&= a + b + c \\
&= 0 \\
&= \mathrm{orthogonal}
\end{aligned}
$$

## Factors

Factors are a data type in R that indicate the values of a variable are categorical. It is important to encode categorical variables in your data frame as factors, as that will tell the statistical functions in R to treat them as such. If you do not do this you can run into unexpected (and incorrect results). You can use the `as_factor()` function (`?forcats::as_factor`) from the **forcats** package to turn a variable in a data frame into a factor.

```{r as_factor}
birthdays <- dplyr::mutate(
  birthdays,
  bdayquarter = forcats::as_factor(bdayquarter)
)
birthdays
```

When you are creating contrast codes for a variable it's important that these codes align with the factor levels of that variable in your data frame. The `levels()` function (`?levels`) can be used to view factor levels of a variable.

```{r levels}
levels(birthdays$bdayquarter)
```

Sometimes the levels of a factor are not in the order you want. The `fct_relevel()` function (`?forcats::fct_relevel`) can be used to relevel factors. Paired with the `mutate()` function (`?dplyr::mutate`) it can be used to relevel factors within a data frame.

```{r}
birthdays_releveled <- dplyr::mutate(
  birthdays,
  bdayquarter = forcats::fct_relevel(
    bdayquarter,
    c("four", "three", "two", "one")
  )
)
levels(birthdays_releveled$bdayquarter)
```

## Setting contrasts

You can set contrasts by creating a series of vectors with your contrast codes, then the `contrasts()` and `cbind()` functions can be used to apply the contrast codes to the variable in your data frame.

```{r set-contrasts}
# Create contrast coding for each hypothesis
contrast_1 <- c(3,-1,-1,-1)
contrast_2 <- c(0, 2,-1,-1)
contrast_3 <- c(0, 0, 1,-1)

# Set contrasts in the birthdays tibble
contrasts(birthdays$bdayquarter) <- cbind(contrast_1, contrast_2, contrast_3)

# View the contrasts after they're set
contrasts(birthdays$bdayquarter)
```

Note that you want your contrasts to be orthogonal. If they are not orthogonal then some additional steps need to be taken to get accurate estimates when you run your contrasts. You do not need to worry about this in lab.

Also note that because the contrast codes are applied to the variable in your data frame, any model you run with that variable will use those contrast codes. This will not be an issue in this course, but it's good to know about. See the links in Learning More for alternative ways of setting contrasts if you are ever in a situation where this behaviour would be an annoyance.

## ANOVA

Once contrasts are set you can fit your ANOVA model, same as we did last week.

```{r aov}
birthdays_model <- aov(awesomeness ~ bdayquarter, data = birthdays)
broom::tidy(birthdays_model)

effectsize::omega_squared(birthdays_model)
```

## Getting contrasts

To calculate the contrast results you can use the same `contrast()` function from the **emmeans** package as last week, but rather than using the argument `method = "pairwise"` to get pairwise contrasts, you will tell the `method` argument the specific contrasts you want to calculate using the contrast code objects specified earlier.

```{r contrast}
birthdays_emm <- emmeans::emmeans(birthdays_model, specs = "bdayquarter")

birthdays_emm_contrast <- emmeans::contrast(
  birthdays_emm,
  method = list("H1" = contrast_1, "H2" = contrast_2, "H3" = contrast_3),
  adjust = "none",
  infer = TRUE
)

birthdays_emm_contrast
```

The formula to calculate the effect size for contrast results is:

$$
d = \frac{2 * \mathrm{estimate}}{\sigma_R * \sum_{k = 1}^k{| c_k |}}
$$

It looks like a lot is going on here, but the equation becomes simpler once you turn it into R code. First you can calculate the sums of the absolute values of the contrast codes for each contrast.

```{r}
contrast_sums <- colSums(abs(contrasts(birthdays$bdayquarter)))
contrast_sums
```

Then you can tidy the emmeans contrast object.

```{r}
birthdays_emm_contrast <- broom::tidy(birthdays_emm)
birthdays_emm_contrast
```

Finally you can plug the formula into the `mutate()` function to compute the correct effect size for each contrast. The `sigma()` function is used to get the residual standard deviation from the fitted ANOVA model.

```{r eff-size}
birthdays_emm_contrast <- dplyr::mutate(
  birthdays_emm_contrast,
  effect.size = (2 * estimate) / (sigma(birthdays_model) * contrast_sums)
)
birthdays_emm_contrast
```

## Learning more

### Contrasts and ANOVA in R

- Deep dive into setting contrasts in R <https://rstudio-pubs-static.s3.amazonaws.com/65059_586f394d8eb84f84b1baaf56ffb6b47f.html>
